---
title: "Misinformation research at the CSS lab"
author: "David Garcia <br> *Graz University of Technology and CSH Vienna*"
date: "28.02.2022"
output:
  xaringan::moon_reader:
    lib_dir: libs 
    css: [xaringan-themer.css, "libs/footer.css"]
    nature:
      beforeInit: ["libs/perc.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
#This block contains the theme configuration for the CSS lab slides style
library(xaringanthemer)
library(showtext)
style_mono_accent(
  base_color = "#5c5c5c",
  text_font_size = "1.5rem",
  header_font_google = google_font("Arial"),
  text_font_google   = google_font("Arial", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


layout: true
<div class="my-footer"><span>David Garcia - Computational Social Science Lab - TU Graz + CSH Vienna</span></div> 

---

# CSS lab and misinformation

> *I don't think I'll do much research on fake news* (David, 2018)


### 1. PRODEMINFO

### 2. Honesty project

### 3. Jigsaw inoculation experiment

### 4. EMOMIS

---

# PRODEMINFO 
> *If everybody always lies to you, the consequence is not that you believe the lies, but rather that nobody believes anything... And with such a people you can then do
what you please.* — Hannah Arendt

- Protecting the Democratic Information Space: Steve Lewandowsky's ERC Advanced grant. TU Graz beneficiary (5-year postdoc, Segun Aroyehun)

- Change in ontologies of truth as the disease behind the symptom of misinformation

- Analysis in 5 languages in news, parliament records, Twitter, etc. Compiation of a corpus of misinformation

- Testing prebunking strategies on Twitter leveraging on the friendship paradox

---



# Honesty project

- Spinoff of PRODEMINFO to test ontology of honesty based on three behaviors:
  - Truth seeking
  - Belief speaking
  - Fostering understanding
  

- Testing predicted shifts towards belief speaking in periods of democratic backsliding
- Funded by the Templeton Foundation (Uni Bristol + TU Graz)
- Analysis of US text (NYT, tweets, COHA, Google books)
- Comparing historical phases with survey results about democratic backsliding and distrust in the media and other institutions

---

# Jigsaw inoculation experiment

- Collaboration with Sander van der Linden, Steve Lewandowsky, Beth Goldberg, Jon Roozenbek, Jana Lasser

- Google Jigsaw provides funding including mass ad campaign on Twitter

- Testing inoculation interventions against fearmongering in the wild

- Evaluating outcomes with observational Twitter data analysis: emotions, link lists from newsguard and other lists (Lazer, De Domenico, etc)

- Pipelines and concept ready, about to launch.

- Side project: extension of misinfo detection on Twitter with ML (Valentin Rupprecht's master thesis)


